### 2016-01-13 papers/CAFe/HIPS-paper/intro.tex
minor tweaks

With the addition of coarrays in the 2008 standard, Fortran is now a \emph{parallel}
language.  Originally called Coarray Fortran (CAF) by its
authors\cite{Numrich:1998:CFP:289918.289920}, parallism in Fortran is similar to MPI in
that all cooperating CAF processes (called \emph{images}) execute a Single Program but with
different (Multiple) Data (SPMD).  Key concepts in CAF are parallel execution, distributed
memory allocation, and remote memory transfer.  Parallel execution and memory allocation
are straightforward as each program is replicated $NP$ times with each program image
allocating its own memory locally.  Special syntax is provided to indicate which arrays
are visible between distributed memory processes and remote memory transfer is indicated
with new array syntax using square brackets, for example,
%\small
\begin{verbatim}
   U(1) = U(1)[3]
\end{verbatim}
%\normalsize
copies memory from the first element of \texttt{U} residing on image number 3 into
the corresponding element of the \emph{local} array \texttt{U}.  The fundamental programming
model is similar to that provided by message passing systems.

@@

In this paper we propose CAFe: Co-Array Fortran Extensions.  These extensions to CAF introduce
relatively minor additions to the existing coarray portion of the Fortran language.  They
include the concept of subimages in order to nest coarray images, semantics for distributed
array declarations, memory placement, remote memory transfer, and remote task execution.  CAFe
builds upon the existing standardized coarray features of Fortran.  We will introduce the
CAFe extensions that are proposed to define their semantics, demonstrate their use in simple
examples, and provide experimental results demonstrating their performance in a prototype implementation.
CAFe is complementary to
previous work extending coarray Fortran\cite{mellor-crummey:2009:caf2,jin:2011:caf2}.

>>>>

With the addition of coarrays in the 2008 standard, Fortran is now a \emph{parallel}
language.  Originally called Coarray Fortran (CAF) by its
authors\cite{Numrich:1998:CFP:289918.289920}, parallelism in Fortran is similar to MPI in
that all cooperating CAF processes (called \emph{images}) execute a Single Program but with
different (Multiple) Data (SPMD).  Key concepts in CAF are parallel execution, distributed
memory allocation, and remote memory transfer.  Parallel execution and memory allocation
are straightforward as each program is replicated $NP$ times with each program image
allocating its own memory locally.  Special syntax is provided to indicate which arrays
are visible between distributed memory processes and remote memory transfer is indicated
with new array syntax using square brackets, for example,
%\small
\begin{verbatim}
   U(1) = U(1)[3]
\end{verbatim}
%\normalsize
copies memory from the first element of \texttt{U} residing on image number 3 into
the corresponding element of the \emph{local} array \texttt{U}.  The fundamental programming
model is similar to that provided by message passing systems.

@@

In this paper we propose CAFe: \emph{Co-Array Fortran Extensions}.  These extensions to CAF introduce
relatively minor additions to the existing coarray portion of the Fortran language.  They
include the concept of subimages in order to nest coarray images, semantics for distributed
array declarations, memory placement, remote memory transfer, and remote task execution.  CAFe
builds upon the existing standardized coarray features of Fortran.  We will introduce the
CAFe extensions that are proposed to define their semantics, demonstrate their use in simple
examples, and provide experimental results demonstrating their performance in a prototype implementation.
CAFe is complementary to
previous work extending coarray Fortran\cite{mellor-crummey:2009:caf2,jin:2011:caf2}.
